Concurrent Programming in Harmony

### 2021-01-19 book/paper.tex
Latest

\autoref{fig:inc} presents a more subtle example that illustrates
non-atomicity.
The program
initializes two shared variables: an integer \textit{count} and
an array \textit{done} with two booleans.
Method \texttt{incrementer} takes a parameter called \textit{self}.
It increments \textit{count} and sets \textit{done}[\textit{self}] to \texttt{True}.
Method \texttt{main} waits until both \textit{done} flags are set to \texttt{True}.
(\textbf{await} $c$ is semantically the same as \textbf{while not} $c$: \textbf{pass}.)
After that, method \texttt{incrementer}
verifies that the value of \textit{count} equals 2.
Like Python, Harmony supports \texttt{assert} statements with two arguments:
the first is a Boolean condition, and the second is a value that is
reported if the condition fails, in this case
the value of \textit{count}.

The program spawns three threads.
The first runs \texttt{incrementer(0)}, the second runs
\texttt{incrementer(1)}, and the last runs \texttt{main()}.
Note that although the threads are \emph{spawned} one at a time,
they will execute concurrently.  It is, for example, quite possible
that \texttt{incrementer(1)} finishes before \texttt{incrementer(0)}
even gets going.
And because Harmony tries every possible execution, it will consider
that particular execution as well.

@@

\begin{figure}
\begin{code}
\begin{verbatim}
code/Up.hny:1 count = 0;
   0 Push 0
   1 Store count
code/Up.hny:2 done = [ False, False ];
   2 Push [False, False]
   3 Store done
code/Up.hny:4 def incrementer(self):
   4 Jump 31
   5 Frame incrementer self
code/Up.hny:5     count = count + 1;
   6 Load count
   7 Push 1
   8 2-ary +
   9 Store count
code/Up.hny:6     done[self] = True;
   10 Push ?done
   11 LoadVar self
   12 Address
   13 Push True
   14 Store
   15 Return
code/Up.hny:8 def main():
   16 Jump 31
   17 Frame main ()
code/Up.hny:9     await all(done);
   18 Load done
   19 1-ary all
   20 JumpCond False 18
code/Up.hny:10     assert count == 2, count;
   21 ReadonlyInc
   22 AtomicInc
   23 Load count
   24 Push 2
   25 2-ary ==
   26 Load count
   27 Assert2
   28 AtomicDec
   29 ReadonlyDec
   30 Return
code/Up.hny:13 spawn incrementer(0);
   31 Push 0
   32 Dup
   33 Push PC(5)
   34 Spawn
\end{verbatim}
\end{code}
\caption{The first part of the HVM bytecode corresponding to \autoref{fig:inc}.}
\label{fig:inccode}
\end{figure}

>>>>

\autoref{fig:inc} presents a more subtle example that illustrates
non-atomicity.
The program initializes two shared variables:
an integer \textit{count} and
an array \textit{done} with two booleans.
The program then spawns three threads.
The first runs \texttt{incrementer(0)}, the second runs
\texttt{incrementer(1)}.

Method \texttt{incrementer} takes a parameter called \textit{self}.
It increments \textit{count} and sets \textit{done}[\textit{self}] to \texttt{True}.
It then waits until the other thread is done.
(\textbf{await} $c$ is semantically the same as \textbf{while not} $c$: \textbf{pass}.)
After that, method \texttt{incrementer}
verifies that the value of \textit{count} equals 2.
Like Python, Harmony supports \texttt{assert} statements with two arguments:
the first is a Boolean condition, and the second is a value that is
reported if the condition fails, in this case
the value of \textit{count}.

Note that although the threads are \emph{spawned} one at a time,
they will execute concurrently.  It is, for example, quite possible
that \texttt{incrementer(1)} finishes before \texttt{incrementer(0)}
even gets going.
And because Harmony tries every possible execution, it will consider
that particular execution as well.

@@

\begin{figure}
\begin{code}
\begin{verbatim}
   0 Frame __init__ ()
code/Up.hny:1 count = 0;
   1 Push 0
   2 Store count
code/Up.hny:2 done = [ False, False ];
   3 Push [False, False]
   4 Store done
code/Up.hny:4 def incrementer(self):
   5 Jump 32
   6 Frame incrementer self
code/Up.hny:5     count = count + 1;
   7 Load count
   8 Push 1
   9 2-ary +
   10 Store count
code/Up.hny:6     done[self] = True;
   11 Push ?done
   12 LoadVar self
   13 Address
   14 Push True
   15 Store
code/Up.hny:7     await done[1 - self];
   16 Push 1
   17 LoadVar self
   18 2-ary -
   19 Load done
   20 Apply
   21 JumpCond False 16
code/Up.hny:8     assert count == 2, count;
   22 ReadonlyInc
   23 AtomicInc
   24 Load count
   25 Push 2
   26 2-ary ==
   27 Load count
   28 Assert2
   29 AtomicDec
   30 ReadonlyDec
   31 Return
\end{verbatim}
\end{code}
\caption{The first part of the HVM bytecode corresponding to \autoref{fig:inc}.}
\label{fig:inccode}
\end{figure}
